---
title: "GDSC Cell Line Data Preprocessing"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libs,  message=FALSE, warning=FALSE} 
library(here)
library(data.table)
library(readxl)
library(dplyr)
library(gtsummary)
library(stringr)
library(ggplot2)
# library(impute)
library(gplots)
```

```{r setup}
plotFolder <- here("images", "GDSC", "02_EDA")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

dataFolder <- here("data", "GDSC")
if(!file.exists(dataFolder)) dir.create(dataFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=TRUE,
  results = 'markup', dev='png', dpi=350, fig.align = "center", fig.path = paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)

set.seed(234)
```

```{r}
make_cdr_heatmap <- function(x, rowColor, colColor, Rowv=T, Colv=T, binary=T, title = "Cancer Drug Response"){
  
  if(binary){
    col_cdr = c("#D9B4D5", "#6BBC47")
    key_show = F
  } else {
    col_cdr = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20)
    key_show = T
  }
  
  
  heatmap_cdr <- heatmap.2(
    x,
    Rowv = Rowv,  # perform hierarchical clustering on rows
    Colv = Colv,  # Perform hierarchical clustering on columns
    col = col_cdr,  # Use the custom binary colors
    dendrogram = "both",
    trace = "none",
    key = key_show,  # Disable the default color key
    RowSideColors = rowColor,
    ColSideColors = colColor,
    labRow = FALSE,  # Hide row names
    labCol = FALSE,  # Hide column names
    main = title,
    lwid = c(0.2, 5),
    lhei = c(1, 5)
  )
  
  # Add legend after creating the heatmap
  legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Sensitivity")
  
  mtext("Cance\n Types", side = 2, line = 0.5, at = 1, las = 2, cex = 0.8)
  
  return(heatmap_cdr)
}
```


# All cancer types
```{r load_data}
c_data <- read.csv(here(dataFolder, "c_data.csv"), row.names = 1)
c_meta <- read.csv(here(dataFolder, "c_meta.csv"))
d_data <- read.csv(here(dataFolder, "d_data.csv"), row.names = 1)
d_meta <- read.csv(here(dataFolder, "Drug_list2022.csv"))

d_meta <- d_meta %>% dplyr::select(Drug.Id, Name, Synonyms, Targets, Target.pathway, PubCHEM, Screening.site) %>% unique()

cdr <- read.csv(here(dataFolder, "cdr.csv"), row.names = 1)
colnames(cdr) <- gsub("^X", "",  colnames(cdr))

cat("Cancer RNAseq data dim: ", dim(c_data))
cat("Drug PCA-based feature data dim: ", dim(d_data))
cat("CDR data dim: ", dim(cdr))
```

```{r cancer_type_counts, fig.height=4, fig.width=6}

c_type_levels = c("ALL", "LAML", "LCML", # leukemia
           "CLL", "DLBC", # lymphoma
           "MM",# blood, myeloma
           "LGG", # Brain Lower Grade Glioma 	胶质瘤, nervous_system
           "GBM", # Glioblastoma multiforme	多形成性胶质细胞瘤, nervous_system
           "MB", # nervous_system
           "NB", # neoblastoma 神经母细胞瘤
           "HNSC", # head & neck 头颈鳞状细胞癌
           "SKCM", # Skin Cutaneous Melanoma	皮肤黑色素瘤
           "BRCA", # breast
           "STAD", # Stomach adenocarcinoma	胃癌
           "ESCA", # Esophageal 食管癌
           "KIRC", # Kidney
           "LIHC", # liver
           "LUAD", "LUSC", # lung, lung_NSCLC
           "SCLC", # lung
           "MESO", # Mesothelioma 间皮瘤, lung
           "COAD/READ", # colon,
           "PAAD", # Pancreatic adenocarcinoma	胰腺癌
           "THCA", # Thyroid carcinoma	甲状腺癌, thyroid
           "OV", # Ovarian serous cystadenocarcinoma	卵巢浆液性囊腺癌, urogenital_system
           "UCEC", # Uterine Corpus Endometrial Carcinoma	子宫内膜癌, urogenital_system
           "PRAD", # Prostate adenocarcinoma	前列腺癌, urogenital_system
           "BLCA", # Bladder, urogenital_system
           "CESC" # Cervical squamous cell carcinoma and endocervical adenocarcinoma 宫颈鳞癌和腺癌, urogenital_system
           ) 

c_meta$cancer_type <- factor(c_meta$cancer_type, levels = c_type_levels)


c_type_color = c("#8B2323", "#FF4040", "#FF7256","#FFD300", "#F6FF00",  # leukemia, lymphoma
                 "#00FFE5",
                 "#5800FF", "#C100FF","#AB82FF", # nervous_system
                 "#CAFF70", "#FFB6C1", "#C1FFC1", "#FFB90F",  "#AEEEEE", "#FF83FA", "#9AFF9A", "#FFBBFF", 
                 "#007BFF",  "#4876FF", "#00BFFF", "#00E5FF", # lung
                 "#FF6A00", "#BFEFFF", "#FF6EB4",
                 "#00FF7B", "#C1FF00","#58FF00",  "#CAFF70", "#00FFB0", # urogenital_system
                 "grey")

names(c_type_color) = c(c_type_levels, "NA")

count_c_type_df <- data.frame("c_type" = factor(names(c_type_color), levels = names(c_type_color)),
                              "Freq"= c(as.numeric(table(c_meta$cancer_type)), sum(is.na(c_meta$cancer_type))),
                              "c_type_color" = c_type_color)

ggplot(count_c_type_df, aes(c_type, Freq, fill = c_type)) +
  geom_col() + 
  scale_x_discrete(limits=rev) + 
  coord_flip() + 
  scale_fill_manual(values = count_c_type_df$c_type_color) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "Cancer type", y = "Frequency")
```

```{r}
map_value_to_color <- function(x) {
  if (is.na(x)) {
    return("darkgray")  # Missing values (NA) will be gray
  } else if (x == "0") {
    return("#E89DA0")   # 0 will be red
  } else if (x == "1") {
    return("#80C1C4") # 1 will be green
  } else {
    return("#B696B6")  # Other values will be blue
  }
}

c_colors <- rep(count_c_type_df$c_type_color, count_c_type_df$Freq)
```

```{r all_cdr_heatmap, fig.height=6, fig.width=10}
set.seed(123)
heatmap_all <- heatmap.2(
  as.matrix(cdr),
  Rowv = FALSE,  # Do not perform hierarchical clustering on rows
  Colv = TRUE,  # Perform hierarchical clustering on columns
  col = c("#D9B4D5", "#6BBC47"),  # Use the custom binary colors
  dendrogram = "both",
  trace = "none",
  key = FALSE,  # Disable the default color key
  RowSideColors = c_colors,  # Row annotation color
  labRow = FALSE,  # Hide row names
  labCol = FALSE,  # Hide column names
  main = "Cancer Drug Response",
  lwid = c(0.2, 5)
)

# Add legend after creating the heatmap
legend(y=1.1, x=-0.075, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Sensitivity")

mtext("Cance\n Types", side = 2, line = 0.5, at = 0.85, las = 2, cex = 0.8) # Add title to RowSideColors
# mtext(text = c_type_levels, side = 2, line = 1, at = seq(0, 1, length.out = length(c_type_levels) + 1) - 0.5/length(c_type_levels), las = 2, cex = 0.8)
```


# Skin
## Input
```{r}
c_types <- c("SKCM")

c_meta_skin <- c_meta[c_meta$cancer_type %in% c_types, ]
c_data_skin <- c_data[rownames(c_data) %in% c_meta_skin$COSMIC_ID, ]
cdr_skin <- cdr[rownames(cdr) %in% c_meta_skin$COSMIC_ID, ]

c_colors_skin <- rep(count_c_type_df$c_type_color[count_c_type_df$c_type %in% c_types], count_c_type_df$Freq[count_c_type_df$c_type %in% c_types])
```

CDR

```{r skin_cdr_heatmap, fig.height=6, fig.width=10}
cdr_tmp = as.matrix(cdr_skin)
    
cdr_row_color = rep('#C1FFC1', nrow(cdr_tmp))
cdr_col_color = rep('white', ncol(cdr_tmp))

cdr_heatmap_hier = make_cdr_heatmap(cdr_tmp, cdr_row_color, cdr_col_color, F, F)
```

Cancer expression data

```{r skin_c_heatmap_hier, fig.width=10, fig.height=6}
heatmap_c <- heatmap.2(as.matrix(c_data_skin),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = c_colors_skin,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```

Drug PCA features
```{r skin_c_heatmap_hier, fig.width=10, fig.height=6}
heatmap_c <- heatmap.2(as.matrix(d_data),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = rep("white", nrow(d_data)),  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```

## Output

```{r}
version_folder = "skin_v7"
cdr_skin_hat <- read.csv(here(dataFolder, version_folder, "GDSC_skin_cdr_hat.csv"), row.names = 1)
c_latent_0 <- read.csv(here(dataFolder, version_folder, "GDSC_skin_c_latent_cluster1.csv"), row.names = 1)
d_latent_0 <- read.csv(here(dataFolder, version_folder, "GDSC_skin_d_latent_cluster1.csv"), row.names = 1)

subcluster = T

if(subcluster) {
  c_latent_1 <- read.csv(here(dataFolder, version_folder, "GDSC_skin_c_latent_cluster2.csv"), row.names = 1)
  d_latent_1 <- read.csv(here(dataFolder, version_folder, "GDSC_skin_d_latent_cluster2.csv"), row.names = 1)
}

c_cluster <- cdr_skin_hat[,c('c_name', 'c_cluster')] %>% unique()
d_cluster <- cdr_skin_hat[,c('d_name', 'd_cluster')] %>% unique()
cd_cluster <-  cdr_skin_hat[,c('c_name', 'd_name', 'cluster')] %>% unique()

```



```{r}
c.names.skin.A <- as.character(c_cluster$c_name[c_cluster$c_cluster == "0"])
if(subcluster) {
  c.names.skin.AB <- as.character(c_cluster$c_name[c_cluster$c_cluster == "0 & 1"])
  c.names.skin.B <- as.character(c_cluster$c_name[c_cluster$c_cluster == "1"])
}
c.names.skin.NA <- as.character(c_cluster$c_name[c_cluster$c_cluster == -1])

d.names.skin.A <- as.character(d_cluster$d_name[d_cluster$d_cluster == "0"])
if(subcluster) {
  d.names.skin.AB <- as.character(d_cluster$d_name[d_cluster$d_cluster == "0 & 1"])
  d.names.skin.B <- as.character(d_cluster$d_name[d_cluster$d_cluster == "1"])
}
d.names.skin.NA <- as.character(d_cluster$d_name[d_cluster$d_cluster == -1])

c.names.skin = c(c.names.skin.A, c.names.skin.AB, c.names.skin.B, c.names.skin.NA)
d.names.skin = c(d.names.skin.A, d.names.skin.AB, d.names.skin.B, d.names.skin.NA)
```



```{r}
library(reshape2)
cdr_hat_skin_wide = dcast(cdr_skin_hat, c_name ~ d_name, value.var = "cdr_hat")

rownames(cdr_hat_skin_wide) = cdr_hat_skin_wide$c_name
cdr_hat_skin_wide = subset(cdr_hat_skin_wide, select = -c_name)

cdr_hat_skin_wide
```


### Summary table
#### Cancer
```{r}
c_meta_skin = c_meta[c_meta$COSMIC_ID %in% c(c.names.skin.A, c.names.skin.AB, c.names.skin.B, c.names.skin.NA), ]
c_meta_skin$cluster = "N/A"
c_meta_skin$cluster[c_meta_skin$COSMIC_ID %in% c.names.skin.A] = "A"
c_meta_skin$cluster[c_meta_skin$COSMIC_ID %in% c.names.skin.B] = "B"
c_meta_skin$cluster[c_meta_skin$COSMIC_ID %in% c.names.skin.AB] = "AB"

library(table1)

table1(~ factor(tissue) + factor(tissue_status) + factor(sample_site) + factor(cancer_type)  | cluster, data=c_meta_skin)


table1(~ factor(gender) + factor(ethnicity) + age_at_sampling | cluster, data=c_meta_skin)
```



```{r}
paste(c(c.names.skin.A, c.names.skin.AB), collapse = ", ")
paste(c(c.names.skin.B, c.names.skin.AB), collapse = ", ")

c_meta_skin_Aonly <- c_meta_skin[c_meta_skin$cluster == "A", -c(1,2)]
write.csv(c_meta_skin_Aonly, here(dataFolder, "GDSC_c_meta_skin_A_only.csv"))

c_meta_skin_Bonly <- c_meta_skin[c_meta_skin$cluster == "B", -c(1,2)]
write.csv(c_meta_skin_Bonly, here(dataFolder, "GDSC_c_meta_skin_B_only.csv"))

c_meta_skin_AandB <- c_meta_skin[c_meta_skin$cluster == "AB", -c(1,2)]
write.csv(c_meta_skin_AandB, here(dataFolder, "GDSC_c_meta_skin_A_and_B.csv"))

c_meta_skin_A <- c_meta_skin[c_meta_skin$cluster %in% c("A", "AB"), -c(1,2)]
write.csv(c_meta_skin_A, here(dataFolder, "GDSC_c_meta_skin_A.csv"))

c_meta_skin_B <- c_meta_skin[c_meta_skin$cluster %in% c("B", "AB"), -c(1,2)]
write.csv(c_meta_skin_B, here(dataFolder, "GDSC_c_meta_skin_B.csv"))
```

### Drug
```{r}
d_meta_skin = d_meta[d_meta$Drug.Id %in% d.names.skin, ]
d_meta_skin$cluster = "N/A"
d_meta_skin$cluster[d_meta_skin$Drug.Id %in% d.names.skin.A] = "A"
d_meta_skin$cluster[d_meta_skin$Drug.Id %in% d.names.skin.B] = "B"
d_meta_skin$cluster[d_meta_skin$Drug.Id %in% d.names.skin.AB] = "AB"

table1(~ factor(Target.pathway) | cluster, data=d_meta_skin)
table1(~factor(Targets) | cluster, data=d_meta_skin)
```

d.names.skin.A:
```{r}
print(d_meta_skin$Name[d_meta_skin$cluster == "A"])

paste(d_meta_skin$Name[d_meta_skin$cluster == "A"], collapse = ", ")
```

d.names.skin.AB:
```{r}
print(d_meta_skin$Name[d_meta_skin$cluster == "AB"])
```

d.names.skin.B:
```{r}
paste(d_meta_skin$Name[d_meta_skin$cluster == "B"], collapse = ", ")
print(d_meta_skin$Name[d_meta_skin$cluster == "B"])
```


```{r}
d_meta_skin_A <- d_meta_skin[d_meta_skin$cluster %in% c("A", "AB"),]
write.csv(d_meta_skin_A, here(dataFolder, "GDSC_d_meta_skin_A.csv"))

d_meta_skin_B <- d_meta_skin[d_meta_skin$cluster %in% c("B", "AB"),]
write.csv(d_meta_skin_B, here(dataFolder, "GDSC_d_meta_skin_B.csv"))
```

#### EDA

```{r skin_cdr_hat_heatmap, fig.height=6, fig.width=10}
cdr_hat_predicts <- cdr_skin_hat[,c('c_name', 'd_name', 'cdr_all')] %>% 
  unique() %>% 
  tidyr::pivot_wider(names_from = d_name, values_from = cdr_all) %>%
  as.data.frame() %>%
  `rownames<-`(.$c_name) %>%
  select(-c_name)


if(subcluster) {
    cdr_tmp = as.matrix(cdr_hat_predicts[c.names.skin, d.names.skin])

    cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("#eca680", length(c.names.skin.AB)),
              rep("#e3716e", length(c.names.skin.B)),  rep("white", length(c.names.skin.NA)))
    cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("#eca680", length(d.names.skin.AB)),
              rep("#e3716e", length(d.names.skin.B)),  rep("white", length(d.names.skin.NA)))
  
} else{
  cdr_tmp = as.matrix(cdr_hat_predicts[c(c.names.skin.A, c.names.skin.NA), c(d.names.skin.A, d.names.skin.NA)])

  cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("white", length(c.names.skin.NA)))
  cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("white", length(d.names.skin.NA)))
  
}

cdr_heatmap_hier = make_cdr_heatmap(cdr_tmp, cdr_row_color, cdr_col_color, F, F, title = "Predicted Cancer Drug Response")
```



```{r skin_cdr_heatmap, fig.height=6, fig.width=10}
if(subcluster) {
    cdr_tmp = as.matrix(cdr_skin[c.names.skin, d.names.skin])

    cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("#eca680", length(c.names.skin.AB)),
              rep("#e3716e", length(c.names.skin.B)),  rep("white", length(c.names.skin.NA)))
    cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("#eca680", length(d.names.skin.AB)),
              rep("#e3716e", length(d.names.skin.B)),  rep("white", length(d.names.skin.NA)))
  
} else{
  cdr_tmp = as.matrix(cdr_skin[c(c.names.skin.A, c.names.skin.NA), c(d.names.skin.A, d.names.skin.NA)])

  cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("white", length(c.names.skin.NA)))
  cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("white", length(d.names.skin.NA)))
  
}

cdr_heatmap_hier = make_cdr_heatmap(cdr_tmp, cdr_row_color, cdr_col_color, F, F)
```


```{r all_cdr_col_skin_order_heatmap, fig.height=6, fig.width=12}
cdr_tmp <- as.matrix(cdr)
cdr_tmp <- cdr_tmp[rev(heatmap_all$rowInd), rev(heatmap_all$colInd)]

cdr_tmp_all_skin <- cdr_tmp

if(min(which(rownames(cdr_tmp_all_skin) %in% c.names.skin)) > 1){
  row_index_1 <- 1:(min(which(rownames(cdr_tmp_all_skin) %in% c.names.skin)) - 1)
} else {
  row_index_1 <- c()
}

row_index <- c(row_index_1, 
               which(rownames(cdr_tmp_all_skin) %in% c.names.skin.A),  
               which(rownames(cdr_tmp_all_skin) %in% c.names.skin.AB),  
               which(rownames(cdr_tmp_all_skin) %in% c.names.skin.B), 
               which(rownames(cdr_tmp_all_skin) %in% c.names.skin.NA), 
               (max(which(rownames(cdr_tmp_all_skin) %in% c.names.skin)) + 1):nrow(cdr_tmp_all_skin))

if(max(which(rownames(cdr_tmp_all_skin) %in% c.names.skin)) < nrow(cdr_tmp_all_skin)){
  row_index <- c(row_index, (max(which(rownames(cdr_tmp_all_skin) %in% c.names.skin)) + 1):nrow(cdr_tmp_all_skin))
}


cdr_tmp_all_skin <- cdr_tmp_all_skin[row_index, c(d.names.skin, colnames(cdr_tmp_all_skin)[! colnames(cdr_tmp_all_skin) %in% d.names.skin])]

cdr_all_row_color = rep("white", nrow(cdr_tmp_all_skin))
cdr_all_col_color = rep("white", ncol(cdr_tmp_all_skin))

cdr_all_row_color[which(rownames(cdr_tmp_all_skin) %in% c.names.skin.A)] = "#f3deb7"
cdr_all_row_color[which(rownames(cdr_tmp_all_skin) %in% c.names.skin.NA)] = "white"

cdr_all_col_color[which(colnames(cdr_tmp_all_skin) %in% d.names.skin.A)] = "#f3deb7"
cdr_all_col_color[which(colnames(cdr_tmp_all_skin) %in% d.names.skin.NA)] = "white"
     
if(subcluster) {
    cdr_all_row_color[which(rownames(cdr_tmp_all_skin) %in% c.names.skin.AB)] = "#eca680"
    cdr_all_row_color[which(rownames(cdr_tmp_all_skin) %in% c.names.skin.B)] = "#e3716e"

    cdr_all_col_color[which(colnames(cdr_tmp_all_skin) %in% d.names.skin.AB)] = "#eca680"
    cdr_all_col_color[which(colnames(cdr_tmp_all_skin) %in% d.names.skin.B)] = "#e3716e"
} 

cdr_heatmap = make_cdr_heatmap(cdr_tmp_all_skin, cdr_all_row_color, cdr_all_col_color, F, F)
```



### c latent
```{r , fig.width=10, fig.height=6}

if(subcluster) {
    c_latent_0_tmp = as.matrix(c_latent_0[c.names.skin, ])

    cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("#eca680", length(c.names.skin.AB)),
              rep("#e3716e", length(c.names.skin.B)),  rep("white", length(c.names.skin.NA)))

} else{
  c_latent_0_tmp = as.matrix(c_latent_0[c(c.names.skin.A, c.names.skin.NA)])

  cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("white", length(c.names.skin.NA)))

}


heatmap.2(
    c_latent_0_tmp,
    Rowv = F,  # perform hierarchical clustering on rows
    Colv = T,  # Perform hierarchical clustering on columns
    col =  colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299), 
    RowSideColors = cdr_row_color,
    trace = "none",  # Do not show trace
    key = T,  # Show color key
    labRow = T,  # Hide row names
    labCol = F,  # Hide column names
)
  
# Add legend after creating the heatmap
legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Embedded Cancer data")


```

```{r , fig.width=10, fig.height=6}

if(subcluster) {
    c_latent_1_tmp = as.matrix(c_latent_1[c.names.skin, ])

    cdr_row_color = c(rep("#f3deb7", length(c.names.skin.A)), rep("#eca680", length(c.names.skin.AB)),
              rep("#e3716e", length(c.names.skin.B)),  rep("white", length(c.names.skin.NA)))


    heatmap.2(
      c_latent_1_tmp,
      Rowv = F,  # perform hierarchical clustering on rows
      Colv = T,  # Perform hierarchical clustering on columns
      col =  colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299), 
      RowSideColors = cdr_row_color,
      trace = "none",  # Do not show trace
      key = T,  # Show color key
      labRow = T,  # Hide row names
      labCol = F,  # Hide column names
  )
    
  # Add legend after creating the heatmap
  legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Embedded Cancer data")

}




```



### d latent
```{r , fig.width=10, fig.height=6}

if(subcluster) {
    d_latent_0_tmp = as.matrix(d_latent_0[d.names.skin, ])

    cdr_row_color =   cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("#eca680", length(d.names.skin.AB)),
              rep("#e3716e", length(d.names.skin.B)),  rep("white", length(d.names.skin.NA)))


} else{
  d_latent_0_tmp = as.matrix(d_latent_0[c(d.names.skin.A, d.names.skin.NA)])

  cdr_row_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("white", length(d.names.skin.NA)))

}


heatmap.2(
    d_latent_0_tmp,
    Rowv = F,  # perform hierarchical clustering on rows
    Colv = T,  # Perform hierarchical clustering on columns
    col =  colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299), 
    RowSideColors = cdr_row_color,
    trace = "none",  # Do not show trace
    key = T,  # Show color key
    labRow = T,  # Hide row names
    labCol = F,  # Hide column names
)
  
# Add legend after creating the heatmap
legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Embedded Cancer data")


```

```{r , fig.width=10, fig.height=6}

if(subcluster) {
    d_latent_1_tmp = as.matrix(d_latent_1[d.names.skin, ])

    cdr_row_color =   cdr_col_color = c(rep("#f3deb7", length(d.names.skin.A)), rep("#eca680", length(d.names.skin.AB)),
              rep("#e3716e", length(d.names.skin.B)),  rep("white", length(d.names.skin.NA)))


    heatmap.2(
      d_latent_1_tmp,
      Rowv = F,  # perform hierarchical clustering on rows
      Colv = T,  # Perform hierarchical clustering on columns
      col =  colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299), 
      RowSideColors = cdr_row_color,
      trace = "none",  # Do not show trace
      key = T,  # Show color key
      labRow = T,  # Hide row names
      labCol = F,  # Hide column names
  )
    
  # Add legend after creating the heatmap
  legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Embedded Cancer data")

}




```


## Other method: BCPlaid
```{r}
library("biclust") 

cdr_bi_skin <- cdr_skin
cdr_bi_skin[is.na(cdr_bi_skin)] = -1
cdr_bi_skin <- as.matrix(cdr_bi_skin)
```


```{r}
# bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), verbose=FALSE)
# 
# heatmapBC(cdr_bi_skin, bicResult)
```
```{r}
# bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), 
#                      background = T,  
#                      background.layer	= -1, verbose=FALSE)
# 
# heatmapBC(cdr_bi_skin, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), background = F, verbose=FALSE)
heatmapBC(cdr_bi_skin, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.8, col.release = 0.8)
heatmapBC(cdr_bi_skin, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.5, col.release = 0.5)
heatmapBC(cdr_bi_skin, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_skin, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.7, col.release = 0.7)

heatmapBC(cdr_bi_skin, bicResult)
```

## Other method: BCCC
```{r}
bicResult <- biclust(cdr_bi_skin, method=BCCC())

heatmapBC(cdr_bi_skin, bicResult)
```

## Other method: BCXmotifs
```{r}
bicResult <- biclust(cdr_bi_skin, method=BCXmotifs())

heatmapBC(cdr_bi_skin, bicResult)
```
## Other method: BCSpectral
```{r}
bicResult <- biclust(cdr_bi_skin, method=BCSpectral())

heatmapBC(cdr_bi_skin, bicResult)
```

## Other method: Bimax Bicluster algorithm
```{r}
bicResult <- biclust(cdr_bi_skin, method=BCBimax())

heatmapBC(cdr_bi_skin, bicResult)
```


## Other method: Questmotif Bicluster algorithm
```{r}
bicResult <- biclust(cdr_bi_skin, method=BCQuest())

heatmapBC(cdr_bi_skin, bicResult)
```



## Other method: Consensus Clustering for cancer subtype identification
```{r  fig.height=6, fig.width=10}
library(CancerSubtypes)

result=ExecuteCC(clusterNum=2,d=c_data_skin,maxK=10,clusterAlg="hc",distance="pearson",title="GBM")

row_idx <- c(which(result$group == 1), which(result$group == 2))
cdr_row_color <- c(rep("lightpink", sum(result$group == 1)), rep("#eca680", sum(result$group == 2)))

cdr_heatmap_skin_subtype = make_cdr_heatmap(as.matrix(cdr_skin[row_idx, ]), cdr_row_color, rep("white", ncol(cdr_skin)), F, T)
```

```{r, fig.width=10, fig.height=6}
heatmap_c_data <- heatmap.2(as.matrix(c_data_skin[row_idx, ]),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = cdr_row_color,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```




## Other method: Consensus Clustering for cancer subtype identification
```{r  fig.height=6, fig.width=10}
library(CancerSubtypes)

result=ExecuteCNMF(c_data_skin,clusterNum=2,nrun=30)

row_idx <- c(which(result$group == 1), which(result$group == 2))
cdr_row_color <- c(rep("lightpink", sum(result$group == 1)), rep("#eca680", sum(result$group == 2)))

cdr_heatmap_skin_subtype = make_cdr_heatmap(as.matrix(cdr_skin[row_idx, ]), cdr_row_color, rep("white", ncol(cdr_skin)), F, T)
```

```{r, fig.width=10, fig.height=6}
heatmap_c_data <- heatmap.2(as.matrix(c_data_skin[row_idx, ]),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = cdr_row_color,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```
