---
title: "GDSC Cell Line Data Preprocessing"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libs,  message=FALSE, warning=FALSE} 
library(here)
library(data.table)
library(readxl)
library(dplyr)
library(gtsummary)
library(stringr)
library(ggplot2)
# library(impute)
library(gplots)
library(tidyr)
library(reshape2)
```

```{r setup}
plotFolder <- here("images", "GDSC", "02_EDA")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

dataFolder <- here("data", "GDSC")
if(!file.exists(dataFolder)) dir.create(dataFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=TRUE,
  results = 'markup', dev='png', dpi=350, fig.align = "center", fig.path = paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)

set.seed(234)
```

```{r}
make_cdr_heatmap <- function(x, rowColor, colColor, Rowv=T, Colv=T, binary=T, title = "Cancer Drug Response"){
  
  if(binary){
    col_cdr = c("#D9B4D5", "#6BBC47")
    key_show = F
  } else {
    col_cdr = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20)
    key_show = T
  }
  
  
  heatmap_cdr <- heatmap.2(
    x,
    Rowv = Rowv,  # perform hierarchical clustering on rows
    Colv = Colv,  # Perform hierarchical clustering on columns
    col = col_cdr,  # Use the custom binary colors
    dendrogram = "both",
    trace = "none",
    key = key_show,  # Disable the default color key
    RowSideColors = rowColor,
    ColSideColors = colColor,
    labRow = FALSE,  # Hide row names
    labCol = FALSE,  # Hide column names
    main = title,
    lwid = c(0.2, 5),
    lhei = c(1, 5)
  )
  
  # Add legend after creating the heatmap
  legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Sensitivity")
  
  mtext("Cance\n Types", side = 2, line = 0.5, at = 1, las = 2, cex = 0.8)
  
  return(heatmap_cdr)
}

make_cdr_heatmap_fillna_binary <- function(x, rowColor, colColor, Rowv=T, Colv=T, binary=T, title = "Cancer Drug Response"){
  
    x[is.na(x)] <- -1
    col_cdr = c("white", "#D9B4D5", "#6BBC47")
    key_show = F
  
  heatmap_cdr <- heatmap.2(
    x,
    Rowv = Rowv,  # perform hierarchical clustering on rows
    Colv = Colv,  # Perform hierarchical clustering on columns
    col = col_cdr,  # Use the custom binary colors
    dendrogram = "both",
    trace = "none",
    key = key_show,  # Disable the default color key
    RowSideColors = rowColor,
    ColSideColors = colColor,
    labRow = FALSE,  # Hide row names
    labCol = FALSE,  # Hide column names
    main = title,
    lwid = c(0.2, 5),
    lhei = c(1, 5)
  )
  
  # Add legend after creating the heatmap
  legend(y=1.15, x=0.0, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Sensitivity")
  
  mtext("Cance\n Types", side = 2, line = 0.5, at = 1, las = 2, cex = 0.8)
  
  return(heatmap_cdr)
}
```


# All cancer types
```{r load_data}
c_data <- read.csv(here(dataFolder, "c_data.csv"), row.names = 1)
c_meta <- read.csv(here(dataFolder, "c_meta.csv"))
d_data <- read.csv(here(dataFolder, "d_data.csv"), row.names = 1)
d_meta <- read.csv(here(dataFolder, "Drug_list2022.csv"))

d_meta <- d_meta %>% dplyr::select(Drug.Id, Name, Synonyms, Targets, Target.pathway, PubCHEM, Screening.site) %>% unique()

cdr <- read.csv(here(dataFolder, "cdr.csv"), row.names = 1)
colnames(cdr) <- gsub("^X", "",  colnames(cdr))

cat("Cancer RNAseq data dim: ", dim(c_data))
cat("Drug PCA-based feature data dim: ", dim(d_data))
cat("CDR data dim: ", dim(cdr))
```

```{r cancer_type_counts, fig.height=4, fig.width=6}

c_type_levels = c("ALL", "LAML", "LCML", # leukemia
           "CLL", "DLBC", # lymphoma
           "MM",# blood, myeloma
           "LGG", # Brain Lower Grade Glioma 	胶质瘤, nervous_system
           "GBM", # Glioblastoma multiforme	多形成性胶质细胞瘤, nervous_system
           "MB", # nervous_system
           "NB", # neoblastoma 神经母细胞瘤
           "HNSC", # head & neck 头颈鳞状细胞癌
           "SKCM", # Skin Cutaneous Melanoma	皮肤黑色素瘤
           "BRCA", # breast
           "STAD", # Stomach adenocarcinoma	胃癌
           "ESCA", # Esophageal 食管癌
           "KIRC", # Kidney
           "LIHC", # liver
           "LUAD", "LUSC", # lung, lung_NSCLC
           "SCLC", # lung
           "MESO", # Mesothelioma 间皮瘤, lung
           "COAD/READ", # colon,
           "PAAD", # Pancreatic adenocarcinoma	胰腺癌
           "THCA", # Thyroid carcinoma	甲状腺癌, thyroid
           "OV", # Ovarian serous cystadenocarcinoma	卵巢浆液性囊腺癌, urogenital_system
           "UCEC", # Uterine Corpus Endometrial Carcinoma	子宫内膜癌, urogenital_system
           "PRAD", # Prostate adenocarcinoma	前列腺癌, urogenital_system
           "BLCA", # Bladder, urogenital_system
           "CESC" # Cervical squamous cell carcinoma and endocervical adenocarcinoma 宫颈鳞癌和腺癌, urogenital_system
           ) 

c_meta$cancer_type <- factor(c_meta$cancer_type, levels = c_type_levels)


c_type_color = c("#8B2323", "#FF4040", "#FF7256","#FFD300", "#F6FF00",  # leukemia, lymphoma
                 "#00FFE5",
                 "#5800FF", "#C100FF","#AB82FF", # nervous_system
                 "#CAFF70", "#FFB6C1", "#C1FFC1", "#FFB90F",  "#AEEEEE", "#FF83FA", "#9AFF9A", "#FFBBFF", 
                 "#007BFF",  "#4876FF", "#00BFFF", "#00E5FF", # lung
                 "#FF6A00", "#BFEFFF", "#FF6EB4",
                 "#00FF7B", "#C1FF00","#58FF00",  "#CAFF70", "#00FFB0", # urogenital_system
                 "grey")

names(c_type_color) = c(c_type_levels, "NA")

count_c_type_df <- data.frame("c_type" = factor(names(c_type_color), levels = names(c_type_color)),
                              "Freq"= c(as.numeric(table(c_meta$cancer_type)), sum(is.na(c_meta$cancer_type))),
                              "c_type_color" = c_type_color)

ggplot(count_c_type_df, aes(c_type, Freq, fill = c_type)) +
  geom_col() + 
  scale_x_discrete(limits=rev) + 
  coord_flip() + 
  scale_fill_manual(values = count_c_type_df$c_type_color) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "Cancer type", y = "Frequency")
```

```{r}
map_value_to_color <- function(x) {
  if (is.na(x)) {
    return("darkgray")  # Missing values (NA) will be gray
  } else if (x == "0") {
    return("#E89DA0")   # 0 will be red
  } else if (x == "1") {
    return("#80C1C4") # 1 will be green
  } else {
    return("#B696B6")  # Other values will be blue
  }
}

c_colors <- rep(count_c_type_df$c_type_color, count_c_type_df$Freq)
```

```{r all_cdr_heatmap, fig.height=6, fig.width=10}
set.seed(123)
heatmap_all <- heatmap.2(
  as.matrix(cdr),
  Rowv = FALSE,  # Do not perform hierarchical clustering on rows
  Colv = TRUE,  # Perform hierarchical clustering on columns
  col = c("#D9B4D5", "#6BBC47"),  # Use the custom binary colors
  dendrogram = "both",
  trace = "none",
  key = FALSE,  # Disable the default color key
  RowSideColors = c_colors,  # Row annotation color
  labRow = FALSE,  # Hide row names
  labCol = FALSE,  # Hide column names
  main = "Cancer Drug Response",
  lwid = c(0.2, 5)
)

# Add legend after creating the heatmap
legend(y=1.1, x=-0.075, xpd=TRUE, legend = c("0", "1"), fill = c("#D9B4D5", "#6BBC47"), bty = "n", cex = 0.8, title = "Sensitivity")

mtext("Cance\n Types", side = 2, line = 0.5, at = 0.85, las = 2, cex = 0.8) # Add title to RowSideColors
# mtext(text = c_type_levels, side = 2, line = 1, at = seq(0, 1, length.out = length(c_type_levels) + 1) - 0.5/length(c_type_levels), las = 2, cex = 0.8)
```


# Leukemia & Lymphoma
## Input
```{r}
c_types <- c("ALL","LAML", "LCML", "CLL", "DLBC")

c_meta_ll <- c_meta[c_meta$cancer_type %in% c_types, ]
c_meta_ll$c_type_updated <- ifelse(c_meta_ll$cancer_type %in% c("ALL","LAML", "LCML", "CLL"), "leukemia", "lymphoma")
c_meta_ll <-  c_meta_ll %>% arrange(c_type_updated)

c_data_ll <- c_data[c_meta_ll$COSMIC_ID, ]
cdr_ll <- cdr[as.character(c_meta_ll$COSMIC_ID), ]

c_colors_ll <- c(rep("#8B2323", sum(c_meta_ll$c_type_updated == "leukemia")), 
                 rep("#F6FF00", sum(c_meta_ll$c_type_updated == "lymphoma")))
```

CDR

```{r skin_cdr_heatmap, fig.height=6, fig.width=10}
cdr_tmp = as.matrix(cdr_ll)
    
cdr_row_color = c_colors_ll
cdr_col_color = rep('white', ncol(cdr_tmp))

cdr_heatmap_hier = make_cdr_heatmap(cdr_tmp, cdr_row_color, cdr_col_color, F, F)
```

```{r skin_cdr_heatmap_clust_fillna, fig.height=6, fig.width=10}
cdr_heatmap_hier = make_cdr_heatmap(cdr_tmp, cdr_row_color, cdr_col_color, T, F)
```

```{r skin_cdr_heatmap_clust_fillna, fig.height=6, fig.width=10}
cdr_heatmap_hier = make_cdr_heatmap_fillna_binary(cdr_tmp, cdr_row_color, cdr_col_color, T, T)
```

Cancer expression data

```{r skin_c_heatmap_hier, fig.width=10, fig.height=6}
heatmap_c <- heatmap.2(as.matrix(c_data_ll),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = F,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299), 
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = c_colors_ll,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```

Drug PCA features
```{r skin_c_heatmap_hier, fig.width=10, fig.height=6}
heatmap_c <- heatmap.2(as.matrix(d_data),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = rep("white", nrow(d_data)),  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Drug PCA Features Data"
                         )
```

## Output

```{r}
version_folder = "ll_6" 

cdr_ll_hat <- read.csv(here(dataFolder, version_folder, 
                            paste0("GDSC_", version_folder, "_cdr_hat.csv")), 
                       row.names = 1)


clust_num <- 4
c_latent_list <- list()
d_latent_list <- list()

for (i in 0:(clust_num-1)){
  c_latent_list[[i+1]] <- read.csv(here(dataFolder, version_folder, paste0("GDSC_", version_folder,"_c_latent_cluster", i, ".csv")), row.names = 1)
  d_latent_list[[i+1]] <- read.csv(here(dataFolder, version_folder, paste0("GDSC_", version_folder,"_d_latent_cluster", i, ".csv")), row.names = 1)
  
  assign(paste0("c_latent_", i), 
         read.csv(here(dataFolder, version_folder, paste0("GDSC_", version_folder,"_c_latent_cluster", i, ".csv")), 
                  row.names = 1))
  assign(paste0("d_latent_", i), 
         read.csv(here(dataFolder, version_folder, paste0("GDSC_", version_folder,"_d_latent_cluster", i, ".csv")), 
                  row.names = 1))
}

cdr_ll_hat$cluster <- gsub("0", "A", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("1", "B", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("2", "C", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("3", "D", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("4", "E", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("5", "F", cdr_ll_hat$cluster)
cdr_ll_hat$cluster[cdr_ll_hat$cluster == "-B"] <- "-1"

cdr_ll_hat$c_cluster <- gsub("0", "A", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("1", "B", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("2", "C", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("3", "D", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("4", "E", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("5", "F", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster[cdr_ll_hat$c_cluster == "-B"] <- "-1"

cdr_ll_hat$d_cluster <- gsub("0", "A", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("1", "B", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("2", "C", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("3", "D", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("4", "E", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("5", "F", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster[cdr_ll_hat$d_cluster == "-B"] <- "-1"

c_cluster <- cdr_ll_hat[,c('c_name', 'c_cluster')] %>% unique()
d_cluster <- cdr_ll_hat[,c('d_name', 'd_cluster')] %>% unique()
cd_cluster <-  cdr_ll_hat[,c('c_name', 'd_name', 'cluster')] %>% unique()
```


```{r}
clust_levels <- row.names(as.matrix(table(cd_cluster$cluster)))
clust_levels <- c(clust_levels[-1], "-1")
cd_cluster$cluster <- factor(cd_cluster$cluster, levels = clust_levels)
table( cd_cluster$cluster, as.integer(cd_cluster$cluster))
cd_cluster$cluster <- as.integer(cd_cluster$cluster)

cd_cluster_wide <- cd_cluster %>% 
  pivot_wider(names_from = d_name, values_from = cluster) %>% 
  as.data.frame()
rownames(cd_cluster_wide) <- cd_cluster_wide$c_name
cd_cluster_wide <- cd_cluster_wide %>% select(-c_name)
```
```{r}
table(d_cluster$d_cluster)
```


```{r}
set.seed(45)
init_heatmap <- heatmap.2(as.matrix(cd_cluster_wide),
                          dendrogram = "both",
                          trace = "none")

cd_cluster_wide_new <- cd_cluster_wide[rev(init_heatmap$rowInd), init_heatmap$colInd]
```


```{r}
#View(cd_cluster_wide_new)

c.names.ll = c(row.names(cd_cluster_wide_new))
d.names.ll = c(colnames(cd_cluster_wide_new))
```



```{r}

cdr_hat_predicts <- cdr_ll_hat[,c('c_name', 'd_name', 'cdr_hat')] %>% 
  unique() %>% 
  tidyr::pivot_wider(names_from = d_name, values_from = cdr_hat) %>%
  as.data.frame() %>%
  `rownames<-`(.$c_name) %>%
  select(-c_name)

cdr_hat_predicts <- cdr_hat_predicts[c.names.ll, d.names.ll]

```



### EDA

```{r skin_cdr_hat_heatmap, fig.height=6, fig.width=10}
heatmap.2(as.matrix(cdr_hat_predicts), 
          Rowv = F, Colv=F, 
          trace='none',
          col = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20))

heatmap.2(as.matrix(cd_cluster_wide[c.names.ll, d.names.ll]),
          Rowv = F, Colv=F,
          trace = "none")

heatmap.2(as.matrix(cdr_ll[c.names.ll, d.names.ll]),
          Rowv = F, Colv=F,
          trace = "none",
          col = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20))


row.names(c_cluster) <- c_cluster$c_name
row.names(d_cluster) <- d_cluster$d_name
c_cluster[c.names.ll, ]
```

# Combine cluster A and B
```{r}
version_folder = "ll_6" 

cdr_ll_hat <- read.csv(here(dataFolder, version_folder, 
                            paste0("GDSC_", version_folder, "_cdr_hat.csv")), 
                       row.names = 1)

cdr_ll_hat$cluster <- gsub("0", "A", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("1", "B", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("2", "C", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("3", "D", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("4", "E", cdr_ll_hat$cluster)
cdr_ll_hat$cluster <- gsub("5", "F", cdr_ll_hat$cluster)
cdr_ll_hat$cluster[cdr_ll_hat$cluster == "-B"] <- "-1"

cdr_ll_hat$c_cluster <- gsub("0", "A", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("1", "B", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("2", "C", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("3", "D", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("4", "E", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster <- gsub("5", "F", cdr_ll_hat$c_cluster)
cdr_ll_hat$c_cluster[cdr_ll_hat$c_cluster == "-B"] <- "-1"

cdr_ll_hat$d_cluster <- gsub("0", "A", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("1", "B", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("2", "C", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("3", "D", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("4", "E", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster <- gsub("5", "F", cdr_ll_hat$d_cluster)
cdr_ll_hat$d_cluster[cdr_ll_hat$d_cluster == "-B"] <- "-1"

cdr_ll_hat$d_cluster[cdr_ll_hat$d_cluster %in% c("A", "A & B", "B", "A & C")] = "B"

cdr_ll_hat$cluster[cdr_ll_hat$cluster %in% c("A", "A & B", "B", "A & C")] = "B"

cdr_ll_hat$c_cluster[cdr_ll_hat$c_cluster  == "A & D"] = "B & D"
cdr_ll_hat$c_cluster[cdr_ll_hat$c_cluster %in% c("A", "A & B", "B", "A & C")] = "B"
cdr_ll_hat$c_cluster = gsub("A & B","B",cdr_ll_hat$c_cluster)


cdr_ll_hat[cdr_ll_hat$c_name == "907274" & cdr_ll_hat$d_name == "1004", "cluster"] = "-1"
cdr_ll_hat[cdr_ll_hat$c_name == "907272" & cdr_ll_hat$d_name == "1004", "cluster"] = "-1"

cdr_ll_hat[cdr_ll_hat$c_name %in% c("907063", "1323913"), "c_cluster"] = "B & C"

cdr_ll_hat$cluster[grepl("B", cdr_ll_hat$c_cluster) & grepl("B", cdr_ll_hat$d_cluster)] = "B"

cdr_ll_hat$cluster <- gsub("B","A", cdr_ll_hat$cluster)
cdr_ll_hat$c_cluster <- gsub("B","A", cdr_ll_hat$c_cluster)
cdr_ll_hat$d_cluster <- gsub("B","A", cdr_ll_hat$d_cluster)

cdr_ll_hat$cluster <- gsub("C","B", cdr_ll_hat$cluster)
cdr_ll_hat$c_cluster <- gsub("C","B", cdr_ll_hat$c_cluster)
cdr_ll_hat$d_cluster <- gsub("C","B", cdr_ll_hat$d_cluster)

cdr_ll_hat$cluster <- gsub("D","C", cdr_ll_hat$cluster)
cdr_ll_hat$c_cluster <- gsub("D","C", cdr_ll_hat$c_cluster)
cdr_ll_hat$d_cluster <- gsub("D","C", cdr_ll_hat$d_cluster)

c_cluster <- cdr_ll_hat[,c('c_name', 'c_cluster')] %>% unique()
d_cluster <- cdr_ll_hat[,c('d_name', 'd_cluster')] %>% unique()
cd_cluster <-  cdr_ll_hat[,c('c_name', 'd_name', 'cluster')] %>% unique()

rownames(c_cluster) <- c_cluster$c_name
rownames(d_cluster) <- d_cluster$d_name

```


```{r}
clust_levels <- row.names(as.matrix(table(cd_cluster$cluster)))
clust_levels <- c(clust_levels[-1], "-1")
cd_cluster$cluster <- factor(cd_cluster$cluster, levels = clust_levels)
table( cd_cluster$cluster, as.integer(cd_cluster$cluster))
cd_cluster$cluster <- as.integer(cd_cluster$cluster)

cd_cluster_wide <- cd_cluster %>% 
  pivot_wider(names_from = d_name, values_from = cluster) %>% 
  as.data.frame()
rownames(cd_cluster_wide) <- cd_cluster_wide$c_name
cd_cluster_wide <- cd_cluster_wide %>% select(-c_name)
```

```{r}
table(d_cluster$d_cluster)
```


```{r}
set.seed(123)
init_heatmap <- heatmap.2(as.matrix(cd_cluster_wide),
                          dendrogram = "both",
                          trace = "none")

cd_cluster_wide_1 <- cd_cluster_wide[rev(init_heatmap$rowInd), init_heatmap$colInd]

cd_cluster_wide_new <- rbind(cd_cluster_wide_1[- which(rowMeans(cd_cluster_wide_1) == 4), ],
                             cd_cluster_wide_1[which(rowMeans(cd_cluster_wide_1) == 4), ])
```


```{r}
c.names.ll = c(row.names(cd_cluster_wide_new))
d.names.ll = c(colnames(cd_cluster_wide_new))
```



```{r}

cdr_hat_predicts <- cdr_ll_hat[,c('c_name', 'd_name', 'cdr_hat')] %>% 
  unique() %>% 
  tidyr::pivot_wider(names_from = d_name, values_from = cdr_hat) %>%
  as.data.frame() %>%
  `rownames<-`(.$c_name) %>%
  select(-c_name)

cdr_hat_predicts <- cdr_hat_predicts[c.names.ll, d.names.ll]

```



### EDA
```{r}
d_cluster <- d_cluster[d.names.ll, ]
c_cluster <- c_cluster[c.names.ll, ]

ColColor <- c(rep("#C79DC9", sum(d_cluster$d_cluster == "A")),
              rep("#EAF7F7", sum(d_cluster$d_cluster == "B")),
              rep("#FDE9F0", sum(d_cluster$d_cluster == "C")),
              rep("white", sum(d_cluster$d_cluster == "-1")))

RowColor <- rep("white", length(c.names.ll))
RowColor[c_cluster$c_cluster == "A"] = "#C79DC9"
RowColor[c_cluster$c_cluster == "A & B"] = "#94D8D8"
RowColor[c_cluster$c_cluster == "B"] = "#EAF7F7"
RowColor[c_cluster$c_cluster == "A & C"] = "#F19ABC"
RowColor[c_cluster$c_cluster == "D"] = "#FDE9F0"
RowColor[c_cluster$c_cluster == "A & B & C"] = "#99A9D1"
```

```{r skin_cdr_hat_heatmap, fig.height=6, fig.width=10}
heatmap.2(as.matrix(cd_cluster_wide[c.names.ll, d.names.ll]),
          Rowv = F, Colv=F,
          ColSideColors = ColColor, RowSideColors = RowColor,
          trace = "none")

heatmap.2(as.matrix(cdr_hat_predicts[c.names.ll, d.names.ll]), 
          Rowv = F, Colv=F,
          ColSideColors = ColColor, 
          RowSideColors = RowColor,
          trace='none',
          col = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20))

heatmap.2(as.matrix(cdr_ll[c.names.ll, d.names.ll]),
          Rowv = F, Colv=F,
          ColSideColors = ColColor, RowSideColors = RowColor,
          trace = "none",
          col = colorRampPalette(c("#D9B4D5",  "white",  "#6BBC47"))(n = 20))


```

```{r all_cdr_col_ll_order_heatmap, fig.height=6, fig.width=12}
cdr_tmp <- as.matrix(cdr)
cdr_tmp <- cdr_tmp[rev(heatmap_all$rowInd), rev(heatmap_all$colInd)]

cdr_tmp <- cdr_tmp[c(c.names.ll, rownames(cdr_tmp)[! rownames(cdr_tmp) %in% c.names.ll]),d.names.ll]

cdr_tmp_all_ll <- cdr_tmp

RowColor_all <- c(RowColor, rep("white", sum(! rownames(cdr_tmp) %in% c.names.ll))) 

cdr_heatmap = make_cdr_heatmap(cdr_tmp_all_ll, RowColor_all, ColColor, F, F)
```


### Summary table
#### Cancer
```{r}
row.names(c_meta) = c_meta$COSMIC_ID
c_meta_ll = c_meta[c.names.ll, ]

row.names(c_cluster) <- c_cluster$c_name
c_meta_ll$cluster = c_cluster[c.names.ll, ]$c_cluster
c_meta_ll$cluster_A <- grepl("A", c_meta_ll$cluster)
c_meta_ll$cluster_B <- grepl("B", c_meta_ll$cluster)
c_meta_ll$cluster_C <- grepl("C", c_meta_ll$cluster)

c_meta_ll_out <- c_meta_ll[, -c(1,2)] %>% select(cluster, cluster_A, cluster_B, cluster_C, everything())

write.csv(c_meta_ll_out, here(dataFolder, "GDSC_c_meta_ll.csv"))
```

```{r}
library(table1)

table1(~ factor(tissue) + factor(tissue_status) + factor(sample_site) + factor(cancer_type)  | cluster, data=c_meta_ll)

table1(~ factor(gender) + factor(ethnicity) + age_at_sampling | cluster, data=c_meta_ll)
```

```{r}
paste0(c_meta_ll_out$COSMIC_ID[c_meta_ll_out$cluster_A], collapse = ", ")
paste0(c_meta_ll_out$COSMIC_ID[c_meta_ll_out$cluster_B], collapse = ", ")
paste0(c_meta_ll_out$COSMIC_ID[c_meta_ll_out$cluster_C], collapse = ", ")
```


### Drug
```{r}
row.names(d_meta) <- d_meta$Drug.Id
d_meta_ll = d_meta[d.names.ll, ]
d_meta_ll$cluster = d_cluster[d.names.ll, ]$d_cluster

d_meta_ll$cluster_A <- grepl("A", d_meta_ll$cluster)
d_meta_ll$cluster_B <- grepl("B", d_meta_ll$cluster)
d_meta_ll$cluster_C <- grepl("C", d_meta_ll$cluster)

d_meta_ll_out <- d_meta_ll %>% select(cluster,  cluster_A, cluster_B, cluster_C, everything())
write.csv(d_meta_ll_out, here(dataFolder, "GDSC_d_meta_ll.csv"))
          
          
table1(~ factor(Target.pathway) | cluster, data=d_meta_ll)
table1(~factor(Targets) | cluster, data=d_meta_ll)
```

```{r}
paste0(d_meta_ll_out$Name[d_meta_ll_out$cluster_A], collapse = ", ")
paste0(d_meta_ll_out$Name[d_meta_ll_out$cluster_B], collapse = ", ")
paste0(d_meta_ll_out$Name[d_meta_ll_out$cluster_C], collapse = ", ")
```


## Other method: BCPlaid
```{r}
library("biclust")

cdr_bi_ll <- cdr_ll
cdr_bi_ll[is.na(cdr_bi_ll)] = -1
cdr_bi_ll <- as.matrix(cdr_bi_ll)
```


```{r}
# bicResult <- biclust(cdr_bi_ll, method=BCPlaid(), verbose=FALSE)
#
# heatmapBC(cdr_bi_ll, bicResult)
```
```{r}
# bicResult <- biclust(cdr_bi_ll, method=BCPlaid(),
#                      background = T,
#                      background.layer	= -1, verbose=FALSE)
#
# heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll, method=BCPlaid(), background = F, verbose=FALSE)
heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.8, col.release = 0.8)
heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.5, col.release = 0.5)
heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll, method=BCPlaid(), background = F, verbose=FALSE,
                     row.release = 0.7, col.release = 0.7)

heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
cdr_bi_ll_tmp <- cdr_bi_ll + 1

bicResult <- biclust(cdr_bi_ll_tmp, method=BCPlaid(), background = 0, verbose=FALSE,
                     row.release = 0.7, col.release = 0.7)

heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll_tmp, method=BCPlaid(), background = 0, verbose=F,
                     row.release = 0.8, col.release = 0.8)

heatmapBC(cdr_bi_ll, bicResult)
```

```{r}
bicResult <- biclust(cdr_bi_ll_tmp, method=BCPlaid(), background = 0, verbose=F,
                     row.release = 0.5, col.release = 0.5)

heatmapBC(cdr_bi_ll, bicResult)
```

## Other method: BCCC
```{r}
bicResult <- biclust(cdr_bi_ll, method=BCCC())

heatmapBC(cdr_bi_ll, bicResult)
```

## Other method: BCXmotifs
```{r}
bicResult <- biclust(cdr_bi_ll, method=BCXmotifs())

heatmapBC(cdr_bi_ll, bicResult)
```
## Other method: BCSpectral
```{r}
bicResult <- biclust(cdr_bi_ll, method=BCSpectral())

heatmapBC(cdr_bi_ll, bicResult)
```

## Other method: Bimax Bicluster algorithm
```{r}
bicResult <- biclust(cdr_bi_ll, method=BCBimax())

heatmapBC(cdr_bi_ll, bicResult)
```


## Other method: Questmotif Bicluster algorithm
```{r}
bicResult <- biclust(cdr_bi_ll, method=BCQuest())

heatmapBC(cdr_bi_ll, bicResult)
```



## Other method: Consensus Clustering for cancer subtype identification
```{r  fig.height=6, fig.width=10}
library(CancerSubtypes)

result=ExecuteCC(clusterNum=2,d=c_data_ll,maxK=10,clusterAlg="hc",distance="pearson",title="GBM")

row_idx <- c(which(result$group == 1), which(result$group == 2))
cdr_row_color <- c(rep("lightpink", sum(result$group == 1)), rep("#eca680", sum(result$group == 2)))

cdr_heatmap_ll_subtype = make_cdr_heatmap(as.matrix(cdr_ll[row_idx, ]), cdr_row_color, rep("white", ncol(cdr_ll)), F, T)
```

```{r, fig.width=10, fig.height=6}
heatmap_c_data <- heatmap.2(as.matrix(c_data_ll[row_idx, ]),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = cdr_row_color,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```




## Other method: Consensus Clustering for cancer subtype identification
```{r  fig.height=6, fig.width=10}
library(CancerSubtypes)

result=ExecuteCNMF(c_data_ll, clusterNum=2, nrun=30) # add the objective 

row_idx <- c(which(result$group == 1), which(result$group == 2))
cdr_row_color <- c(rep("lightpink", sum(result$group == 1)), rep("#eca680", sum(result$group == 2)))

cdr_heatmap_ll_subtype = make_cdr_heatmap(as.matrix(cdr_ll[row_idx, ]), cdr_row_color, rep("white", ncol(cdr_ll)), F, T)
```

```{r, fig.width=10, fig.height=6}
heatmap_c_data <- heatmap.2(as.matrix(c_data_ll[row_idx, ]),  # Exclude non-numeric columns
                         Rowv = F,  # Do not perform hierarchical clustering on rows
                         Colv = T,  # perform hierarchical clustering on columns
                         col = colorRampPalette(c("blue", "purple", "white", "orange", "red"))(n = 299),  # Apply custom colors to heatmap
                         dendrogram = "both",  # show dendrograms
                         trace = "none",  # Do not show trace
                         key = T,  # Show color key
                         RowSideColors = cdr_row_color,  # Row annotation color
                         labRow = T,  # Hide row names
                         labCol = F,  # Hide column names
                         main = "Cancer preprocessed RNAseq Data"
                         )
```
