---
title: "Simulations of cancer data,  drug data, and corrresponding cancer drug response data"
author: "Seraphina Shi"
date: "2023-03-17"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(gplots)
library(reshape)

plotFolder <- here("images", "simulations")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

dataFolder <- here("data", "simulations")
if(!file.exists(dataFolder)) dir.create(dataFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=TRUE,
  results = 'markup', dev='png', dpi=150, fig.align = "center",
  fig.path=paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)
```

```{r}
GDSC_folder <- here("data", "GDSC")
```

In this R file, we simulate multiple synthetic samples based on lung cancer cell
lines. 

# Original data
Load GDSC lung cancer cell lines 
```{r}
# cancer data
RNAseq_rpkm <- read.csv(file=here(GDSC_folder, "GDSC_lung_RNAseq_rpkm.csv"), row.names = 1, header= TRUE)
names(RNAseq_rpkm) <- gsub("X", "", names(RNAseq_rpkm))

cl_meta <- read.csv(file=here(GDSC_folder, "GDSC_CellLine_lung_annotation.csv"), header=TRUE)
cl_meta <- cl_meta[cl_meta$COSMIC_ID %in% names(RNAseq_rpkm) & (!is.na(cl_meta$COSMIC_ID)), ]

# drug data

# Cancer drug response data
CDR_auc_bi <- read.csv(file=here(GDSC_folder, "GDSC_DrugResponse_binarized.csv"), row.names = 1, header= TRUE)
names(CDR_auc_bi) <- gsub("X", "", names(CDR_auc_bi))
```

# Simulate data 1 (GDSC, NSCLC)

## Cancer 
```{r fig.width=6, fig.height=7}
#---------------------------------------------------
# 0. check cancer types
#---------------------------------------------------
table(cl_meta$cancer_type)

#---------------------------------------------------
# 1. focus on NSCLC and SCLC 
#---------------------------------------------------
NSCLC_COSMIC <- cl_meta$COSMIC_ID[cl_meta$cancer_type == "NSCLC"]
NSCLC_COSMIC <- NSCLC_COSMIC[!is.na(NSCLC_COSMIC)]

NSCLC_RNAseq <- RNAseq_rpkm %>% select(all_of(as.character(NSCLC_COSMIC)))
dim(NSCLC_RNAseq) # [1] 5703   76

heatmap.2(as.matrix(NSCLC_RNAseq),
          trace="none",
          col = rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)) ,
          labRow=FALSE,
          main = "Heatmap of NSCLC cell line data",
          key.title = "Color Key\n& Histogram",
          key.xlab=NA,
          keysize=1
          )
```

```{r}
#---------------------------------------------------
# 2.1 select some consistently highly expressed or lowly expressed genes for each cancer type
#. 2.2.   select several genes with high expression variability for each cancer type
#. 2.3 keep these genes in the simulated data
#---------------------------------------------------
set.seed(123)
row_means <- apply(NSCLC_RNAseq, 1, mean)
row_vars <- apply(NSCLC_RNAseq, 1, var)
summary(row_means)
summary(row_vars)

NSCLC_consistently_high_genes <- row.names(NSCLC_RNAseq)[which(row_vars < 0.35 & (row_means > 0.4 ))] 
NSCLC_consistently_low_genes <- row.names(NSCLC_RNAseq)[which(row_vars < 0.35 & (row_means < -0.4 ))] 
NSCLC_consistently_med_genes <- row.names(NSCLC_RNAseq)[which(row_vars < 0.35 & (-0.1 < row_means & row_means < 0.1))] 

NSCLC_high_var_expressed_genes <- row.names(NSCLC_RNAseq)[which(row_vars > 2.3 )]


selected_genes <- unique(c(NSCLC_consistently_high_genes, NSCLC_consistently_low_genes,
                               NSCLC_consistently_med_genes,  NSCLC_high_var_expressed_genes))
NSCLC_random_genes <- sample(row.names(NSCLC_RNAseq)[! row.names(NSCLC_RNAseq) %in% selected_genes], size=50)

c(length(NSCLC_consistently_high_genes), length(NSCLC_consistently_low_genes), length(NSCLC_consistently_med_genes), length(NSCLC_high_var_expressed_genes), length(NSCLC_random_genes))

all_selected_genes <- unique(c(NSCLC_consistently_high_genes, NSCLC_consistently_low_genes,
                               NSCLC_consistently_med_genes,  NSCLC_high_var_expressed_genes,
                               NSCLC_random_genes))
length(all_selected_genes) 


NSCLC_RNAseq_s <- NSCLC_RNAseq %>% 
                    t() %>% as.data.frame() %>% 
                    select(all_of(all_selected_genes)) %>% 
                    t() %>% as.data.frame()

dim(NSCLC_RNAseq_s)
```

```{r fig.width=6, fig.height=7}
#---------------------------------------------------
# 4. separate the cancers into two groups
#---------------------------------------------------
cat_genes <- NSCLC_high_var_expressed_genes[1:10]
library(corrplot)
corrplot(cor(t(NSCLC_RNAseq_s[NSCLC_high_var_expressed_genes,])))

apply(NSCLC_RNAseq_s[c("SIDG05348", "SIDG05982", "SIDG09035", "SIDG09036", "SIDG09100", "SIDG10617", "SIDG37641"),], 1, mean)
sep <- NSCLC_RNAseq["SIDG05348",] < 0.5 & NSCLC_RNAseq["SIDG09035",] < 0.5 & 
  NSCLC_RNAseq["SIDG09036",] < 0.5 & NSCLC_RNAseq["SIDG09100",] < 0.5 & NSCLC_RNAseq["SIDG10617",] < 0.5

grp1 <- names(NSCLC_RNAseq_s)[which(sep)]
grp2 <- names(NSCLC_RNAseq_s)[which(!sep)]

grp1_RNAseq <- NSCLC_RNAseq_s %>% 
  select(all_of(as.character(grp1))) 

grp2_RNAseq <- NSCLC_RNAseq_s %>% 
  select(all_of(as.character(grp2))) 


RNAseq_simu <- cbind(grp1_RNAseq, grp2_RNAseq)
RNAseq_meta_simu <- cbind(C_ID = colnames(RNAseq_simu), 
                          C_type = c(rep("grp1", length(grp1)), rep("grp2", length(grp2))))


heatmap.2(as.matrix(RNAseq_simu),
          trace="none",
          col = rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)) ,
          Colv=FALSE,
          Rowv = FALSE,
          labRow=FALSE,
          main = "Heatmap of cancer cell line data",
          ColSideColors = c(rep("#00BFC4",length(grp1)), rep("#F8766D", length(grp2))),
          key.title = "Color Key\n& Histogram",
          key.xlab=NA,
          keysize=1
          )
legend(y=1.05, x=.12, xpd=TRUE,
       legend = c("group1","group2"),
       col = c("#00BFC4", "#F8766D"),
       lty= 1, lwd = 5, cex=.7
       )
```


## Drug

Since drug data is complicated and can be expressed through multiple ways, (smile strings, molecular fingerprints, and molecular graphs); we start with fingerprints that are easier to simulate and easier to create the cancer-drug-response functions. 

```{r}
set.seed(123)
p <- 100
# 10 drugs to start with
drug_simu_func <- function(n, p, k_fp_idx){
  fp <- matrix(nrow = n, ncol = p)
  for(i in 1:p){
    if(i %in% k_fp_idx){
      fp[,i] = rbinom(n=n, size = 1, prob = 0.95)
    } else {
      fp[, i] = rbinom(n=n, size = 1, prob = 0.5)
    }
  }
  return(fp)
}

n_d = 5
p = 150

d_fp <- rbind(drug_simu_func(n_d, p, 1:20),
              drug_simu_func(n_d, p, 50:70))

colnames(d_fp) <- paste0("f", 1:p)
rownames(d_fp) <- c(paste0("d_g1_", 1:n_d), paste0("d_g2_", 1:n_d))

d_fp <- t(d_fp)


heatmap.2(as.matrix(d_fp),
          trace="none",
          # col = rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)) ,
          Colv=FALSE,
          Rowv = FALSE,
          labRow=FALSE,
          main = "Heatmap of drug fingureprints data",
          ColSideColors = c(rep("#00BFC4",n_d), rep("#F8766D", n_d)),
          key.title = "Color Key\n& Histogram",
          key.xlab=NA,
          keysize=1
          )
legend(y=1.05, x=.12, xpd=TRUE,
       legend = c("group1","group2"),
       col = c("#00BFC4", "#F8766D"),
       lty= 1, lwd = 5, cex=.7
       )
```
    

## Cancer Drug Response
```{r}
n_c =  ncol(RNAseq_simu)
n_d = ncol(d_fp)/2
cdr <- matrix(0, nrow = n_c, ncol = n_d*2)
cdr[1:length(grp1), 1:n_d] = 1
cdr[(length(grp1)+1):n_c, (n_d+1):(n_d*2)] = 1
colnames(cdr) = colnames(d_fp)
rownames(cdr) = colnames(RNAseq_simu)

heatmap.2(t(as.matrix(cdr)),
          trace="none",
          # col = rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)) ,
          Colv=FALSE,
          Rowv = FALSE,
          main = "Heatmap of drug fingureprints data",
          ColSideColors = c(rep("#00BFC4",length(grp1)), rep("#F8766D", length(grp2))),
          key.title = "Color Key\n& Histogram",
          key.xlab=NA,
          keysize=1
          )
legend(y=1.05, x=.12, xpd=TRUE,
       legend = c("group1","group2"),
       col = c("#00BFC4", "#F8766D"),
       lty= 1, lwd = 5, cex=.7
       )

```


```{r}
write.csv(RNAseq_simu, here(dataFolder, "simu1_RNAseq.csv"))
write.csv(RNAseq_meta_simu, here(dataFolder, "simu1_RNAseq_meta.csv"))
write.csv(d_fp, here(dataFolder, "simu1_d_fp.csv"))
write.csv(cdr, here(dataFolder, "simu1_cdr.csv"))
```




